\sys: Light-touch Data-Parallel Shell Processing

### 2020-05-26 pldi/draft.tex
Typos and minor fixes

The \unix shell is an environment---often interactive---for composing scripts written in a plethora of programming languages.
This language-agnosticism, coupled with \unix's toolbox philosophy~\cite{mcilroy1978unix}, makes the shell the primary choice for specifying succinct and simple pipelines that target data processing, system orchestration, and other automation tasks.
% Pipe-centric shell commands are parallel (and to some extent, functional) programs that can natively enjoy significant parallelism on a multicore machine.
Unfortunately, parallelizing such pipelines requires significant effort shared between two different programmer groups. % it's more like roles personas or "wearing hats"

The first group is \emph{command developers}, responsible for developing individual commands such as \ttt{sort}, \ttt{uniq}, and \ttt{jq}.
These developers usually work in a single programming language, leveraging its abstractions to provide parallelism whenever possible.
As they have no visibility into the command's uses, they expose a plethora of ad-hoc command-specific flags such as \ttt{-t}, \ttt{-}\ttt{-parallel}, \ttt{-p}, and \ttt{-j} % , \ttt{CPUs}, and \ttt{NUM\_THREADS}~\cite{}.
\cite{pasetto2011comparative, mcilroy1993engineering, stallman1991gnu}.

The second group is \emph{shell users}, who use POSIX shell constructs to combine multiple such commands from many languages into their scripts and are thus left with only a few options for incorporating parallelism.
% \kk{I don't really like the reference to such a later section. Can we remove it?}
% \nv{We have to say this is not just ``an opinion'', but that we give a scientific treatment/full coverage in the related work section. How far away should the sections be so that you're comfortable referencing to?}
% If versed into parallelism, 
% Users versed in computing can apply 
% these tools are command-unaware thus at risk of breaking program semantics,
% require manual changes to the code to make parallelism explicit,
% at times leverage command flags like the ones mentioned earlier---but each command 
% oblivious? agnostic?  
One option is manual tools such as GNU \ttt{parallel}~\cite{Tange2011a}, \ttt{ts}~\cite{tsp}, \ttt{qsub}~\cite{gentzsch2001sun}, \textsc{SLURM}~\cite{yoo2003slurm};
  these tools are either command-unaware, and thus at risk of breaking program semantics, or too coarse-grained, and thus only capable of exploiting parallelism at the level of entire scripts rather than individual components.
% Tools will attempt to parallelize anything 
% use the composition abstractions of the \unix shell, 
% and are often prototypical in nature---focusing on quickly prototyping a solution for the task at hand rather building an optimal solution to a recurring pattern.
% short lived, changing quickly
% (ii) ad-hoc flags provided by command developers and mentioned above.
Another option is to use shell primitives such as \ttt{\&}, \ttt{wait}, \ttt{for} to explicitly induce parallelism;
 these come at a cost of manual effort to split inputs, rewrite scripts, and orchestrate execution---an expensive and error-prone process.
% re-writing of the entire script from scratch; unfortunately, this process is expensive and can introduce new bugs, cascading changes, and divergence from legacy functionality---but also foregoes all the productivity benefits of the original script.
% The fact that a programmer often belongs to both groups at the same time---as composed scripts can themselves form a command used in other scrips---compounds the challenge.
To top it off, all these options assume a good understanding of parallelism;
  users whose domain of expertise falls outside computing---routinely writing shell scripts to accomplish their tasks---are left without options.
%  (see~\S\ref{related} for a full discussion).
% \nv{I still don't like ``for a full discussion''---raises too many expectations, and we don't have that much more to say}

@@

To address cold-start issues, \sys comes with a library of parallelizability semantics for commands in POSIX and GNU Coreutils.
These large classes of commands serve as the shell's standard library, expected to be found pervasively. % in scripts found out in the wild.
The study that led to their characterization also informed \sys's annotation and transformation components.\footnote{
  In fact, the scope of the study is broader, incorporating also commands outside POSIX and GNU Coreutils such as \ttt{curl} and \ttt{pandoc}.
}

@@

% In the ongoing debate on climate change,
Suppose an environmental scientist    % looking for a non-CS expert---journalist might be too low
  wants to get a quick sense of the trends in the maximum temperature across the U.S. over the past ten years.
As the National Oceanic and Atmospheric Administration (NOAA) has made historic temperature data publicly available~\cite{noaa}, 
  answering this question is only a matter of a simple data-processing pipeline.

The shell script in Fig.~\ref{fig:example} starts by pulling the yearly index files and filtering out URLs that are not part of the compressed dataset.
It then downloads and decompresses each file in the remaining set, extracts the values that indicate the temperature, and filters out bogus inputs marked as \ttt{999}.
It then calculates the maximum yearly temperature by sorting the values and picking the top element.
Finally, it matches each maximum value with the appropriate year in order to print the result.
% 
The effort required to write this script is astonishingly low:\footnote{
  Some effort is required to understand NOAA's weather format, but this is true for any program processing any dataset.
}
  its data-processing core amounts to 12 stages and, when expressed as a single line, is only 165 characters long.
Its succinctness is deceiving---this program is no toy:
  a Java program implementing only the last four pipeline stages requires 137LoC~\cite[\S2.1]{hadoop:15}.
To enable such a succinct program composition, \unix incorporates several features.

@@

Command flags, used pervasively in \unix, are configuration options that the command's developer has decided to expose to its users to improve the command's general applicability.
For example, by omitting  \ttt{sort}'s \ttt{r} flag that enables reverse sorting, the user can easily get the minimum temperature.
The shell does not have any visibility into these flags; 
  after it expands special characters such as \ttt{\textasciitilde{}} and \ttt{*}, it leaves parsing and evaluation entirely up to individual commands.

@@

\heading{Implementation} %
Independently of the semantics of individual commands and the shell language, the \unix environment has its own set of quirks.
Thus, any attempt to orchestrate paralllel execution will have to take care of several challenges related to task parallelism, deadlock prevention, and runtime performance.
For example, forked processes piping their combined results to Fig.~\ref{fig:example}'s \ttt{head} might not receive a \ttt{PIPE} signal if \ttt{head} exits prior to opening all pipes.
% A final challenge has to do with practical issues, as even manual parallelization of a given shell script is far from trivial.
% Parallelizing a script without reimplementing all of its commands, requires primitives to compose existing commands to achieve parallelization.
%  Designing correct and efficient compositions is particularly challenging in the context of \unix, and requires addressing issues related to scheduling, synchronization, and deadlock prevention. 
Moreover, specific commands such as \ttt{sort} and \ttt{uniq} require specialized data aggregators so as to be correctly parallelized.
%% Developing correct, high-performance combiners for certain classes of commands such as \ttt{sort} and \ttt{uniq} is non trivial.
%% Moreover, the \unix environment has its set of quirks % about opening streams, deadling with FIFOs,
%%   and any attempt at parallelization will have to take care of these challenges---related to task parallelism, deadlock prevention, and runtime performance.

@@

\heading{Scripts}
To maintain sequential semantics, \sys first analyzes a script to identify \emph{parallelizable regions} that contain commands that are candidates for parallelization.
This analysis is guided by the script structure: % to identify parallelization candidates:
  some constructs (\eg \ttt{\&}, \ttt{|}) expose task parallelism; others (\eg \ttt{\&\&}, \ttt{||}) enforce synchronization.
% \sys first needs to understand the POSIX constructs used to compose larger programs, indicating different levels of candidacy for pararllelization.
%% To assign a candidacy for parallelization, \sys breaks POSIX constructs into a few broad classes:
%%  (i) the parallel-composition operator (\ttt{\&}) exposes task parallelism,
%%  (ii) special commands such as \ttt{xargs} and \ttt{tee} expose data parallelism,
%%  (iii) the pipe operator (\ttt{|}) is a candidate for both task and data parallelism, and
%%  (iv)  other composition operators such as \ttt{\&\&}, \ttt{||}, and \ttt{;} are barrier candidates.
% Since the shell representation only exposes coarse grained parallelization opportunities, 
\sys then converts each parallelizable region to a \emph{dataflow graph} (DGF), 
%% Provided information commands and POSIX constructs, \sys needs to analyze the given script and identify program fragments that are candidates for parallelization.
%% To achieve this, it converts sections of the script that are amenable to parallelization into a data-flow graph (DFG) representation.
% Second, \sys defines a dataflow graph model that is suitable for the semantics of the POSIX shell and proposes a set of parallelization optimizations that preserve the behavior of the dataflow graph.
  a flexible representation that enables a series of local transformations to expose data parallelism, converting the graph into its parallel equivalent.
%% , in which DFGs indicate the parallelizable regions.
Further transformations compile the DFG back to a shell script that uses POSIX constructs to guide parallelism explicitly while aiming at preserving the semantics of the sequential program.

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.49\textwidth]{\detokenize{./figs/dish_example2.pdf}}
\caption{
  \textbf{\sys's $2\times$-parallel output for fragment of Fig.~\ref{fig:example}.}
  \sys orchestrates the script's parallel execution through named pipes, parallel operators and careful placement of combiners.
}
\vspace{-15pt}
\label{fig:example2}
\end{figure}

@@

\heading{Parallelizable Pure Commands}
The second class, \pur, contains commands that respect functional purity---\ie same outputs for same inputs---but maintain internal state across their entire pass.
The details of this state and its propagation during element processing affect their parallelizability characteristics.
Some commands are easy to parallelize, because they maintain trivial state and are commutative---\eg \ttt{wc} simply maintains a counter.
Other commands such as \ttt{sort}, maintain more complex invariants that have to be taken into account when merging partial results.
% 
% \begin{figure}[t]
% \centering
% \includegraphics[width=0.20\textwidth]{\detokenize{./figs/dish_classes.pdf}}
% \caption{
%   \textbf{Hierarchy of parallelizability classes.}
%   Parallelizability classes form a hierarchy from the easiest to parallelize (left) to the most difficult (right).
%   \nv{The figure is incorrect---it's actually a lattice.}
% }
% \vspace{-15pt}
% \label{fig:hierarchy}
% \end{figure}

@@

This is the largest class, for two reasons.
First, it includes commands related to the file-system---a central abstraction of the \unix design and philosophy~\cite{unix}.
In fact, \unix uses the file-system as a proxy to several file-unrelated operations such as access control and device driving. % is grafted upon the file-system.
Second, it contains commands that do not consume input or do not produce output---and thus not amenable to data parallelism.
For example, \ttt{date}, \ttt{uname}, and \ttt{finger} are all commands interfacing with kernel- or hardware-generated information and do not consume any input from user programs.
  %A significant part of these commands (a ratio of 21:4) only \emph{read} values, and usually values that are never written by user code.

@@

\heading{Key Concerns}
\sys's annotations focus on three crucial concerns:
  (i) a command's parallelizability class,
  (ii) for commands that support multiple inputs, the characteristics of input consumption, and
  (iii) for commands with multiple flags (options), the handling of competing sets of flags.
As the first concern was discussed extensively in the previous section, we now focus on the latter two.

@@

\heading{Example Annotations}
An annotation record contains a set of clauses, each one identified by a predicate on the command options.
Each clause maps to an assignment for the command, indicating its class and the sequence of its inputs and outputs.
Annotations treat flag arguments differently from file arguments, by checking if they start with a \ttt{-}.
% \footnote{
%   Note that the single character \ttt{-} is not an option as it is most often used to designate that a command reads from its standard input.
%   \nv{
%     There are options of of the form --f "x; y", and BSD options do not take \ttt{-} (like \ttt{ps aux}).
%     I am also not sure how important this last sentence is, as it did not answer a question I had---I expected arguments to be anything: if soneome wants to put ``pizza'' the DSL should not disallow it.
%   }
% }
The annotation grammar is shown in Appendix~\ref{alg}.

@@

\noindent
The first two clauses describe \ttt{comm} invocations with options \ttt{-1} (or \ttt{-2}) and \ttt{-3}.
As the first (second) and third columns are suppressed, \ttt{comm} can be considered in \sta by regarding its first file argument as a static ``configuration'' input;
% NV: I don't think this is correct, because the small input might be the other argument;  you could even have `comm - -`
  this input accompanies every replicated instance of the command executing in parallel.
The third clause is the general case, which includes \ttt{comm} invocations without any options, is captured by the third clause.
There, \ttt{comm} is classified as \pur as it needs to % wait until it 
  read both inputs completely before it can output its results.
Its inputs are its first and second file arguments (\ie excluding any options), and its output is produced in \ttt{stdout}.

@@

\heading{Custom Aggregators}
For commands in \sta, the annotations are enough to enable parallelization:
  commands are applied to parts of their input in parallel, and their outputs are simply concatenated.
%% commands either fall under \sta, having straightforward aggregators, or fall under a subset of \pur that can be parallelized using divide-and-conquer techniques~\sx{optimizer}. \kk{pure commands are what we have to provide aggregators for, so this is not totally valid.}
%% Unfortunately, there are commands that
%% require less obvious tactics---after all, \unix commands are Turing-complete programs and failing to plan for this generality would inhibit \sys's usefulness. \kk{I would propose deleting this last sentence. we have already separated commands into classes, so this sentence makes it seem like the classes don't help at all.}
%
To support the parallelization of arbitrary commands in \pur, \sys allows supplying custom \emph{map} and \emph{aggregate} functions.
In line with the \unix philosophy, these functions can be written in any language as long as they conform to a few invariants:
  (i) \emph{map} is in \sta and \emph{aggregate} is in \pur,
  (ii) \emph{map} can consume (or extend) the output of the original command and  \emph{aggregate} can consume (and combine) the results of multiple \emph{map} invocations, and
  (iii) their composition produces the same output with the original command.
If these invariants are met, after DFG analysis~\sx{ir} \sys can replace occurrences of these commands with their \emph{map} and \emph{aggregate} equivalents, enabling their full parallelization.

@@

% replacing FIFOs with files is
A solution would be to replace FIFOs with files, a central \unix abstraction, simulating pipes of arbitrary buffering (Fig. \ref{fig:eager}b).
Aside from severe performance implications, naive replacement can lead to subtle race conditions as a consumer might reach {\sc EOF} before a producer.
Alternatively, consumers could wait for producers to complete before opening the file for reading (Fig. \ref{fig:eager}c);
however, this would insert artificial barriers impeding task-based parallelism and wasting disk.
% \unix does not provide an easy way out.

@@

\heading{Dangling FIFOs and Zombie Producers}
%
%% \TODO{(This problem is a nice followup for the one above) Problem:
%%   Dataflow graph termination. Illustration: Show the example of head
%%   deadlock and explain that this is inherent behaviour in the shell.
%%   (Solution: emptying pipelines or terminating processes) In order for
%%   things to terminate gracefully, dish wraps all commands in order to
%%   always consume all of their input by sending it down to
%%   /dev/null. Warning: This means that all commands in the pipeline
%%   must terminate gracefully if all of their output is consumed (if a
%%   process dies unexpectedly, noone will consume its input. We leave
%%   addressing that and other monitoring questions for future work). We
%%   evaluate the overhead of emptying all outputs in section X.}
%
% \nv{I think all of these titles should be the problem/challenge, not the solution}
% A process writing to a pipe needs to be notified when its consumer exits early before consuming all its output.
Under normal operation, a command exits after it has produced and sent all its results to its output channel.
If the channel is a pipe and its reader exits early, the command is notified to stop writing early.
In \unix, this is achieved by an out-of-band error mechanism: the
operating system delivers a \ttt{PIPE} signal to the producer,
notifying it that the pipe's consumer has exited.  This is different
from errors for other system calls
%% return an error
and unusual compared to non-\unix systems\footnote{For example,
  Windows indicates errors for \ttt{WriteFile} using its return
  code---similar to \ttt{DeleteFile} and other Win32 functions.}
primarily because pipes and pipelines are at the heart of \unix.
% write errors are communicated differently 
Unfortunately though, if a pipe has not been opened for writing yet% ---possible for named pipes---
  \unix cannot signal this condition.
% \kk{I think that it is also possible for anonymous, e.g. \ttt{cat f1 - | head -n 1}}
% \nv{TODO! Will think about this later}
Consider the following script:
% emitted code from \sys's earlier versions:
%% \kk{Proposed alternative: ``Consider the following shell script:''}
\begin{lstlisting}[language=sh, numbers=none]
  mkfifo fifo1 fifo2
  cat file-chunk1.txt > fifo1 &
  cat file-chunk2.txt > fifo2 &
  cat fifo1 fifo2 | head -n 1 & wait
\end{lstlisting}
\noindent
In the code above, \ttt{head} exits early causing the last \ttt{cat} to exit before opening \ttt{fifo2}.
As a result, the second \ttt{cat} never receives a \ttt{PIPE} signal that its consumer exited---after all, \ttt{fifo2} never even had a consumer.
This, in turn, leaves it unable to make progress, as it is both blocked and unaware of its consumer exiting.
Coupled with \ttt{wait} at the end, the entire snippet reaches a deadlock.

@@

The vast majority requires no effort other than invoking \sys;
  the annotation for the remaining \todo{6} commands ($<3\%$) amounts to a single record.
% 
In terms of correctness, \sys's results over multi-GB inputs are identical to the sequential for all benchmarks.
Scripts feature ample opportunities for breaking semantics, which \sys avoids.
In one case, \sys avoids transforming a cycle (\ie a fragment whose input and output identifiers are the same).
%in another, it preserves the semantics that another parallelization command breaks.

@@

\heading{Setup}
\sys was run on 512GB of memory and 64 physical $\times$ 2.1GHz Intel Xeon E5-2683 cores, Debian 4.9.144-3.1, GNU Core-utils 8.30-3, GNU Bash 5.0.3(1)-, Python 3.7.3, and OCaml 4.05.0---without any special configuration in hardware or software.
Except as otherwise noted,
  (i) all pipelines are set to (initially) read from and (finally) write to the file-system,
  (ii) \ttt{curl} fetches data from a different physical host on the same network connected by 1Gbps links.
% A more detailed description of the setup and experiments, including reproducible scripts, is included in the accompanying (anonymized) online repository.

@@

\heading{Take-aways}
\sys operates on programs with annotated commands outside the POSIX/GNU subsets and leads to notable speedups even when the original program features significant task-based parallelism.
\sys does not risk breaking semantics, thus does not improve the first component. % this will not change

>>>>

The \unix shell is an environment---often interactive---for composing scripts written in a plethora of programming languages.
This language-agnosticism, coupled with \unix's toolbox philosophy~\cite{mcilroy1978unix}, makes the shell the primary choice for specifying succinct and simple pipelines for data processing, system orchestration, and other automation tasks.
% Pipe-centric shell commands are parallel (and to some extent, functional) programs that can natively enjoy significant parallelism on a multicore machine.
Unfortunately, parallelizing such pipelines requires significant effort shared between two different programmer groups. % it's more like roles personas or "wearing hats"

The first group is \emph{command developers}, responsible for implementing individual commands such as \ttt{sort}, \ttt{uniq}, and \ttt{jq}.
These developers usually work in a single programming language, leveraging its abstractions to provide parallelism whenever possible.
As they have no visibility into the command's uses, they expose a plethora of ad-hoc command-specific flags such as \ttt{-t}, \ttt{-}\ttt{-parallel}, \ttt{-p}, and \ttt{-j} % , \ttt{CPUs}, and \ttt{NUM\_THREADS}~\cite{}.
\cite{pasetto2011comparative, mcilroy1993engineering, stallman1991gnu}.

The second group is \emph{shell users}, who use POSIX shell constructs to combine multiple such commands from many languages into their scripts and are thus left with only a few options for incorporating parallelism.
% \kk{I don't really like the reference to such a later section. Can we remove it?}
% \nv{We have to say this is not just ``an opinion'', but that we give a scientific treatment/full coverage in the related work section. How far away should the sections be so that you're comfortable referencing to?}
% If versed into parallelism, 
% Users versed in computing can apply 
% these tools are command-unaware thus at risk of breaking program semantics,
% require manual changes to the code to make parallelism explicit,
% at times leverage command flags like the ones mentioned earlier---but each command 
% oblivious? agnostic?  
One option is manual tools such as GNU \ttt{parallel}~\cite{Tange2011a}, \ttt{ts}~\cite{tsp}, \ttt{qsub}~\cite{gentzsch2001sun}, \textsc{SLURM}~\cite{yoo2003slurm};
  these tools are either command-unaware, and thus at risk of breaking program semantics, or too coarse-grained, and thus only capable of exploiting parallelism at the level of entire scripts rather than individual components.
% Tools will attempt to parallelize anything 
% use the composition abstractions of the \unix shell, 
% and are often prototypical in nature---focusing on quickly prototyping a solution for the task at hand rather building an optimal solution to a recurring pattern.
% short lived, changing quickly
% (ii) ad-hoc flags provided by command developers and mentioned above.
Another option is to use shell primitives (such as \ttt{\&}, \ttt{wait}, \ttt{for}) to explicitly induce parallelism;
 these come at a cost of manual effort to split inputs, rewrite scripts, and orchestrate execution---an expensive and error-prone process.
% re-writing of the entire script from scratch; unfortunately, this process is expensive and can introduce new bugs, cascading changes, and divergence from legacy functionality---but also foregoes all the productivity benefits of the original script.
% The fact that a programmer often belongs to both groups at the same time---as composed scripts can themselves form a command used in other scrips---compounds the challenge.
To top it off, all these options assume a good understanding of parallelism;
  users whose domain of expertise falls outside computing---routinely writing shell scripts to accomplish their tasks---are left without options.
%  (see~\S\ref{related} for a full discussion).
% \nv{I still don't like ``for a full discussion''---raises too many expectations, and we don't have that much more to say}

@@

To address cold-start issues, \sys comes with a library of parallelizability semantics for commands in POSIX and GNU Coreutils.
These large classes of commands serve as the shell's standard library, expected to be used pervasively. % in scripts found out in the wild.
The study that led to their characterization also informed \sys's annotation and transformation components.\footnote{
  In fact, the scope of the study is broader, incorporating also commands outside POSIX and GNU Coreutils such as \ttt{curl} and \ttt{pandoc}.
}

@@

% In the ongoing debate on climate change,
Suppose an environmental scientist    % looking for a non-CS expert---journalist might be too low
  wants to get a quick sense of trends in the maximum temperature across the U.S. over the past ten years.
As the National Oceanic and Atmospheric Administration (NOAA) has made historic temperature data publicly available~\cite{noaa}, 
  answering this question is only a matter of a simple data-processing pipeline.

The shell script in Fig.~\ref{fig:example} starts by pulling the yearly index files and filtering out URLs that are not part of the compressed dataset.
It then downloads and decompresses each file in the remaining set, extracts the values that indicate the temperature, and filters out bogus inputs marked as \ttt{999}.
It then calculates the maximum yearly temperature by sorting the values and picking the top element.
Finally, it matches each maximum value with the appropriate year in order to print the result.
% 
The effort required to write this script is astonishingly low:\footnote{
  Some effort is required to understand NOAA's weather format, but this is true for any program processing any dataset.
}
  its data-processing core amounts to 12 stages and, when expressed as a single line, is only 165 characters long.
Its succinctness is deceiving---this program is no toy:
  a Java program implementing only the last four pipeline stages requires 137~LoC~\cite[\S2.1]{hadoop:15}.
To enable such a succinct program composition, \unix incorporates several features.

@@

Command flags, used pervasively in \unix, are configuration options that the command's developer has decided to expose to its users to improve the command's general applicability.
For example, by omitting  \ttt{sort}'s \ttt{-r} flag that enables reverse sorting, the user can easily get the minimum temperature.
The shell does not have any visibility into these flags; 
  after it expands special characters such as \ttt{\textasciitilde{}} and \ttt{*}, it leaves parsing and evaluation entirely up to individual commands.

@@

\heading{Implementation} %
Independently of the semantics of individual commands and the shell language, the \unix environment has its own set of quirks.
Thus, any attempt to orchestrate parallel execution will have to take care of several challenges related to task parallelism, deadlock prevention, and runtime performance.
For example, forked processes piping their combined results to Fig.~\ref{fig:example}'s \ttt{head} might not receive a \ttt{PIPE} signal if \ttt{head} exits prior to opening all pipes.
% A final challenge has to do with practical issues, as even manual parallelization of a given shell script is far from trivial.
% Parallelizing a script without reimplementing all of its commands, requires primitives to compose existing commands to achieve parallelization.
%  Designing correct and efficient compositions is particularly challenging in the context of \unix, and requires addressing issues related to scheduling, synchronization, and deadlock prevention. 
Moreover, several commands such as \ttt{sort} and \ttt{uniq} require specialized data aggregators so as to be correctly parallelized.
%% Developing correct, high-performance combiners for certain classes of commands such as \ttt{sort} and \ttt{uniq} is non trivial.
%% Moreover, the \unix environment has its set of quirks % about opening streams, deadling with FIFOs,
%%   and any attempt at parallelization will have to take care of these challenges---related to task parallelism, deadlock prevention, and runtime performance.

@@

\heading{Scripts}
To maintain sequential semantics, \sys first analyzes a script to identify \emph{parallelizable regions} that contain commands that are candidates for parallelization.
This analysis is guided by the script structure: % to identify parallelization candidates:
  some constructs (\eg \ttt{\&}, \ttt{|}) expose task parallelism; others (\eg \ttt{\&\&}, \ttt{||}) enforce synchronization.
% \sys first needs to understand the POSIX constructs used to compose larger programs, indicating different levels of candidacy for pararllelization.
%% To assign a candidacy for parallelization, \sys breaks POSIX constructs into a few broad classes:
%%  (i) the parallel-composition operator (\ttt{\&}) exposes task parallelism,
%%  (ii) special commands such as \ttt{xargs} and \ttt{tee} expose data parallelism,
%%  (iii) the pipe operator (\ttt{|}) is a candidate for both task and data parallelism, and
%%  (iv)  other composition operators such as \ttt{\&\&}, \ttt{||}, and \ttt{;} are barrier candidates.
% Since the shell representation only exposes coarse grained parallelization opportunities, 
\sys then converts each parallelizable region to a \emph{dataflow graph} (DFG), 
%% Provided information commands and POSIX constructs, \sys needs to analyze the given script and identify program fragments that are candidates for parallelization.
%% To achieve this, it converts sections of the script that are amenable to parallelization into a data-flow graph (DFG) representation.
% Second, \sys defines a dataflow graph model that is suitable for the semantics of the POSIX shell and proposes a set of parallelization optimizations that preserve the behavior of the dataflow graph.
  a flexible representation that enables a series of local transformations to expose data parallelism, converting the graph into its parallel equivalent.
%% , in which DFGs indicate the parallelizable regions.
Further transformations compile the DFG back to a shell script that uses POSIX constructs to guide parallelism explicitly while aiming at preserving the semantics of the sequential program.

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.49\textwidth]{\detokenize{./figs/dish_example2.pdf}}
\caption{
  \textbf{\sys's $2\times$-parallel output for fragment of Fig.~\ref{fig:example}.}
  \sys orchestrates the script's parallel execution through named pipes, parallel operators and careful placement of combiners. \TODO{Correct Dish!}
}
\vspace{-15pt}
\label{fig:example2}
\end{figure}

@@

\heading{Parallelizable Pure Commands}
The second class, \pur, contains commands that respect functional purity---\ie same outputs for same inputs---but maintain internal state across their entire pass.
The details of this state and its propagation during element processing affect their parallelizability characteristics.
Some commands are easy to parallelize, because they maintain trivial state and are commutative---\eg \ttt{wc} simply maintains a counter.
Other commands, such as \ttt{sort}, maintain more complex invariants that have to be taken into account when merging partial results.
% 
% \begin{figure}[t]
% \centering
% \includegraphics[width=0.20\textwidth]{\detokenize{./figs/dish_classes.pdf}}
% \caption{
%   \textbf{Hierarchy of parallelizability classes.}
%   Parallelizability classes form a hierarchy from the easiest to parallelize (left) to the most difficult (right).
%   \nv{The figure is incorrect---it's actually a lattice.}
% }
% \vspace{-15pt}
% \label{fig:hierarchy}
% \end{figure}

@@

This is the largest class, for two reasons.
First, it includes commands related to the file-system---a central abstraction of the \unix design and philosophy~\cite{unix}.
In fact, \unix uses the file-system as a proxy to several file-unrelated operations such as access control and device driving. % is grafted upon the file-system.
Second, it contains commands that do not consume input or do not produce output---and thus are not amenable to data parallelism.
For example, \ttt{date}, \ttt{uname}, and \ttt{finger} are all commands interfacing with kernel- or hardware-generated information and do not consume any input from user programs.
  %A significant part of these commands (a ratio of 21:4) only \emph{read} values, and usually values that are never written by user code.

@@

\heading{Key Concerns}
\sys's annotations focus on four crucial concerns:
  (i) a command's parallelizability class,
  (ii) a command's inputs and outputs,
  (ii) for commands that support multiple inputs, the characteristics of input consumption, and
  (iii) for commands with multiple flags (options), the handling of competing sets of flags.
As the first concern was discussed extensively in the previous section, we now focus on the latter three.

@@

\heading{Example Annotations}
The parallelizability properties of each command are written in an annotation record that
%% An annotation record
contains a set of clauses, each one identified by a predicate on the command options.
Each clause maps to an assignment for the command, indicating its class and the sequence of its inputs and outputs.
Annotations treat flag arguments differently from file arguments, by checking if they start with a \ttt{-}.
% \footnote{
%   Note that the single character \ttt{-} is not an option as it is most often used to designate that a command reads from its standard input.
%   \nv{
%     There are options of of the form --f "x; y", and BSD options do not take \ttt{-} (like \ttt{ps aux}).
%     I am also not sure how important this last sentence is, as it did not answer a question I had---I expected arguments to be anything: if soneome wants to put ``pizza'' the DSL should not disallow it.
%   }
% }
The complete annotation grammar is shown in Appendix~\ref{alg}.

@@

\noindent
The first two clauses describe \ttt{comm} invocations with options \ttt{-1} (or \ttt{-2}) and \ttt{-3}.
As the first (second) and third columns are suppressed, \ttt{comm} can be considered in \sta by regarding its first file argument as a static ``configuration'' input;
% NV: I don't think this is correct, because the small input might be the other argument;  you could even have `comm - -`
  this input accompanies every replicated instance of the command executing in parallel.
The third clause is the general case, which includes \ttt{comm} invocations without any options, and is captured by the third clause.
There, \ttt{comm} is classified as \pur as it needs to % wait until it 
  read both inputs completely before it can output its results.
Its inputs are its first and second file arguments (\ie excluding any options), and its output is produced in \ttt{stdout}.

@@

\heading{Custom Aggregators}
For commands in \sta, the annotations are enough to enable parallelization:
  commands are applied to parts of their input in parallel, and their outputs are simply concatenated.
%% commands either fall under \sta, having straightforward aggregators, or fall under a subset of \pur that can be parallelized using divide-and-conquer techniques~\sx{optimizer}. \kk{pure commands are what we have to provide aggregators for, so this is not totally valid.}
%% Unfortunately, there are commands that
%% require less obvious tactics---after all, \unix commands are Turing-complete programs and failing to plan for this generality would inhibit \sys's usefulness. \kk{I would propose deleting this last sentence. we have already separated commands into classes, so this sentence makes it seem like the classes don't help at all.}
%
To support the parallelization of arbitrary commands in \pur, \sys allows supplying custom \emph{map} and \emph{aggregate} functions.
In line with the \unix philosophy, these functions can be written in any language as long as they conform to a few invariants:
  (i) \emph{map} is in \sta and \emph{aggregate} is in \pur,
  (ii) \emph{map} can consume (or extend) the output of the original command and  \emph{aggregate} can consume (and combine) the results of multiple \emph{map} invocations, and
  (iii) their composition produces the same output as the original command.
If these invariants are met, after DFG analysis~\sx{ir} \sys can replace occurrences of these commands with their \emph{map} and \emph{aggregate} equivalents, enabling their full parallelization.

@@

% replacing FIFOs with files is
One might be tempted to replace FIFOs with files, a central \unix abstraction, simulating pipes of arbitrary buffering (Fig. \ref{fig:eager}b).
Aside from severe performance implications, naive replacement can lead to subtle race conditions as a consumer might reach {\sc EOF} before a producer.
Alternatively, consumers could wait for producers to complete before opening the file for reading (Fig. \ref{fig:eager}c);
however, this would insert artificial barriers impeding task-based parallelism and wasting disk.
% \unix does not provide an easy way out.

@@

\heading{Dangling FIFOs and Zombie Producers}
%
%% \TODO{(This problem is a nice followup for the one above) Problem:
%%   Dataflow graph termination. Illustration: Show the example of head
%%   deadlock and explain that this is inherent behaviour in the shell.
%%   (Solution: emptying pipelines or terminating processes) In order for
%%   things to terminate gracefully, dish wraps all commands in order to
%%   always consume all of their input by sending it down to
%%   /dev/null. Warning: This means that all commands in the pipeline
%%   must terminate gracefully if all of their output is consumed (if a
%%   process dies unexpectedly, noone will consume its input. We leave
%%   addressing that and other monitoring questions for future work). We
%%   evaluate the overhead of emptying all outputs in section X.}
%
% \nv{I think all of these titles should be the problem/challenge, not the solution}
% A process writing to a pipe needs to be notified when its consumer exits early before consuming all its output.
Under normal operation, a command exits after it has produced and sent all its results to its output channel.
If the channel is a pipe and its reader exits early, the command is notified to stop writing early.
In \unix, this is achieved by an out-of-band error mechanism: the
operating system delivers a \ttt{PIPE} signal to the producer,
notifying it that the pipe's consumer has exited.  This is different
from errors for other system calls
%% return an error
and unusual compared to non-\unix systems\footnote{For example,
  Windows indicates errors for \ttt{WriteFile} using its return
  code---similar to \ttt{DeleteFile} and other Win32 functions.}
primarily because pipes and pipelines are at the heart of \unix.
% write errors are communicated differently 
Unfortunately though, if a pipe has not been opened for writing yet
% ---possible for named pipes---
  \unix cannot signal this condition.
% \kk{I think that it is also possible for anonymous, e.g. \ttt{cat f1 - | head -n 1}}
% \nv{TODO! Will think about this later}
Consider the following script:
% emitted code from \sys's earlier versions:
%% \kk{Proposed alternative: ``Consider the following shell script:''}
\begin{lstlisting}[language=sh, numbers=none]
  mkfifo fifo1 fifo2
  cat file-chunk1.txt > fifo1 &
  cat file-chunk2.txt > fifo2 &
  cat fifo1 fifo2 | head -n 1 & wait
\end{lstlisting}
\noindent
In the code above, \ttt{head} exits early causing the last \ttt{cat} to exit before opening \ttt{fifo2}.
As a result, the second \ttt{cat} never receives a \ttt{PIPE} signal that its consumer exited---after all, \ttt{fifo2} never even had a consumer.
This, in turn, leaves it unable to make progress, as it is both blocked and unaware of its consumer exiting.
Coupled with \ttt{wait} at the end, the entire snippet reaches a deadlock.

@@

The vast majority requires no effort other than invoking \sys;
  the annotation for the remaining \todo{6} commands ($<3\%$) amounts to a single record.
% 
In terms of correctness, \sys's results over multi-GB inputs are identical to the sequential for all benchmarks.
Scripts feature ample opportunities for breaking semantics, which \sys avoids.
%% In one case, \sys avoids transforming a cycle (\ie a fragment whose input and output identifiers are the same).
%in another, it preserves the semantics that another parallelization command breaks.

@@

\heading{Setup}
\sys was run on 512GB of memory and 64 physical $\times$ 2.1GHz Intel Xeon E5-2683 cores, Debian 4.9.144-3.1, GNU Coreutils 8.30-3, GNU Bash 5.0.3(1), and Python 3.7.3---without any special configuration in hardware or software.
%% and OCaml 4.05.0 %% this was here for the parser that we don't mention atm
Except as otherwise noted,
  (i) all pipelines are set to (initially) read from and (finally) write to the file-system,
  (ii) \ttt{curl} fetches data from a different physical host on the same network connected by 1Gbps links.
% A more detailed description of the setup and experiments, including reproducible scripts, is included in the accompanying (anonymized) online repository.

@@

\heading{Take-aways}
\sys operates on programs with annotated commands outside the POSIX/GNU subsets and leads to notable speedups even when the original program features significant task-based parallelism.
%% \sys does not risk breaking semantics, thus does not improve the first component. % this will not change
